---
title: "plotly"
author: "Sumit Kant"
date: "15 September 2018"
output:
  ioslides_presentation:
    css: custom_style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## What is plotly

Plotly is a web application for creating and sharing data
visualizations. Plotly can work with several programming languages
and applications including R, Python, and Microsoft Excel. We're
going to concentrate on creating different graphs with Plotly and
sharing those graphs.



## Installing Plotly

```{r, echo=FALSE, message=FALSE}
# install.packages("plotly")
library(plotly)
```
If you want to share your visualizations on https://plot.ly/ you
should make an account on their site


## Basic Scatterplot

```{r cars, echo = TRUE, message=FALSE}
library(plotly)
plot_ly(mtcars, x = ~wt, y = ~mpg, type="scatter", height=350, width=780)
```

## Scatterplot with color
Higher cylinder counts have higher weight and generaly low miles per gallon
```{r, message=FALSE, echo=TRUE}
plot_ly(mtcars, x = ~wt, y = ~mpg, type = "scatter", color = ~factor(cyl),
        height = 350, width = 780)
```

## Scatterplot - Continuous color
Coloring the scatterplot with displacement as the continous variable.
As shown in the graph, lower disp values tends to have higher miles per gallon and lower weight.
```{r, message=FALSE, echo=T, warning=FALSE}
plot_ly(mtcars, x = ~wt, y = ~mpg, mode = "markers", color = ~disp,
        height = 350, width = 780)
```

## Scatterplot - Sizing
Coloring the scatterplot with displacement as the continous variable.
As shown in the graph, lower disp values tends to have higher miles per gallon and lower weight.
```{r, message=FALSE, echo=T, warning=FALSE}
plot_ly(mtcars, x = ~wt, y = ~mpg, mode = "markers", size = ~hp,
        color = ~factor(cyl), height = 350, width = 780)
```

## 3D Scatterplot

```{r, message=FALSE, echo=T, warning=FALSE}
set.seed(2016-07-21)
temp <- rnorm(100, mean=30, sd =5)
pressue <- rnorm(100)
dtime <- 1:100
plot_ly(x = temp, y = pressue, z = dtime, type="scatter3d", 
        mode = "markers",color = temp, width = 780, height=320)
```

## Line Graph 

The default graph in plotly is a Line Chart. E.g.: Plotting miles travelled over the years in air with airmiles dataset

```{r, message=FALSE, echo=T, warning=FALSE}
data("airmiles")
plot_ly(x = time(airmiles), y = airmiles, mode="lines",
        width = 780, height=350)
```

## Multi-line Graph
Stock market data representation is the most common form of multi-line graphs
```{r, message=FALSE, echo=T, warning=FALSE}
library(tidyr)
library(dplyr)
data("EuStockMarkets")
stocks <- as.data.frame(EuStockMarkets) %>%
  gather(index, price) %>%
  mutate(time = rep(time(EuStockMarkets), 4))
plot_ly(stocks, x = ~time, y = ~price, color = ~index, type = "scatter", 
        mode = "lines", width = 780, height=300)
```


## Histogram
Requires one variable to visualize histogram
```{r, message=FALSE, echo=T, warning=FALSE}
data("iris")
plot_ly(x = iris$Sepal.Length, type = 'histogram',
        width = 780, height=370)
```

## Boxplot
```{r, message=FALSE, echo=T, warning=FALSE}
plot_ly(iris, y = ~Petal.Length, color = ~Species, type = 'box',
        width = 780, height=370)
```


## HeatMap
Plotly takes a matrix as an input and produces a raster image for heatmap
```{r, message=FALSE, echo=T, warning=FALSE}
terrain <- matrix(rnorm(100*100), nrow= 100, ncol=100)
plot_ly(z = terrain, type = 'heatmap', width = 780, height=370)
```

## 3D Surface
Same as heatmap, only type="surface"
```{r, message=FALSE, echo=T, warning=FALSE}
terrain <- matrix(sort(rnorm(100*100)), nrow= 100, ncol=100)
plot_ly(z = terrain, type = 'surface', width = 780, height=370)
```

## Chloropeth Maps (1/2)
Choropleth maps illustrate data across geographic areas by shading regions with different colors. Choropleth maps are easy to make with Plotly though they require more setup compared to other Plotly graphics.
```{r, message=FALSE, echo=T, warning=FALSE, eval =F}
# Create data frame
state_pop <- data.frame(State = state.abb, Pop = as.vector(state.x77[,1]))
# Create hover text
state_pop$hover <- with(state_pop, paste(State, '<br>', "Population:", Pop))
# Make state borders white
borders <- list(color = toRGB("red"))
# Set up some mapping options
map_options <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)
plot_ly(z = ~state_pop$Pop, text = ~state_pop$hover, locations = ~state_pop$State,
        type = 'choropleth', locationmode = 'USA-states', color = state_pop$Pop, 
        colors = 'Blues', marker = list(line = borders), width =750, height=370) %>%
  layout(title = 'US Population in 1975', geo = map_options)
```

## Chloropeth Maps (2/2)
```{r, message=FALSE, echo=F, warning=FALSE}
plot_ly(z = ~state_pop$Pop, text = ~state_pop$hover, locations = ~state_pop$State,
        type = 'choropleth', locationmode = 'USA-states', color = state_pop$Pop, 
        colors = 'Blues', marker = list(line = borders), width =750, height=370) %>%
  layout(title = 'US Population in 1975', geo = map_options)

```